Supplemental Materials for ``Universals of word order result from optimization of grammars for efficient communication''

### 2019-09-13 writeup/pnas-word-order-si.tex

Greenberg's original correlations/implications % ?? TODO

\cite{justeson1990explanation} conducted a log-linear analysis on typological judgments of 147 languages, constructing an undirected graphical model.
For the relations examined here, beyond correlations with verb-object order, they found additional correlations between G-N, N-Adp and between Rel-N, N-Adp.
With their model, they found no evidence for a direct correlation V-O, G-N.

@@

Results are shown in Figure~\ref{fig:toy-parseability}. For all parameter regimes, the correlating grammars have better parseability than the anti-correlating grammars. This is especially the case for grammars with high $p_\text{branching}$.

>>>>

We followed \citet{dryer1992greenbergian} in treating Greenberg's correlations as pairwise correlations with verb-object order.
While Greenberg's original study \cite{greenberg1963universals} also formalized most of these as correlations with verb-object order, a few were formalized as correlations between other relations that are only indirectly related to verb-object order (e.g., Universal 22 linking the position of the standard of comparison to the order of adpositions).

\citet{justeson1990explanation} conducted a log-linear analysis on typological judgments of 147 languages, constructing an undirected graphical model.
%While they found evidence for correlations similar to those found by \citet{dryer1992greenbergian}, they found that some correlations are me
For the relations examined here, beyond correlations with verb-object order, they found additional correlations between (1) the noun-genitive and adposition-noun dependencies, and (2) between the noun-relative clause and adposition-noun dependencies, \emph{beyond} the correlation mediated through the individual correlations with the verb-object dependency.
%With their model, they found no evidence for a direct correlation V-O, G-N beyond that mediated through other correlations.
We ran the same Bayesian logistic mixed-effects analysis for these two correlations.
Results are shown in Figure~\ref{tab:corr-regression-implication}.
Both correlations are very strongly supported by grammars optimized for efficiency.
Future research might apply the graphical model approach of \citet{justeson1990explanation} to optimized grammars, and test whether grammar optimization predicts not just the presence of correlations, the structure their method found on typological data.

\begin{table}
\small{
\begin{center}
\begin{tabular}{|ll||l|lll|llll|ll|llllll}
\hline
	& &Prevalence & Mean & SD & $p(\beta \leq 0)$  \\
\hline\hline
	G-N (nmod) & N-Adp (lifted\_case)  & 0.919 & 4.482 & 1.058 & $<$ \num{1e-4} \\
	Rel-N (acl) & N-Adp (lifted\_case) & 0.898 & 4.653 & 1.286 & $<$ \num{1e-4} \\
\hline
\end{tabular}
\end{center}
}
	\caption{Detailed results for the two correlations found by \citet{justeson1990explanation} that do not involve the verb-object dependency, for grammars optimized for efficiency. Both correlations are strongly supported by optimized grammars, holding in about 90\% of optimized grammars. Compare Table~\ref{tab:corr-regression}.
}\label{tab:corr-regression-implication}
\end{table}

@@

Results are shown in Figure~\ref{fig:toy-parseability}. For almost all parameter regimes, the correlating grammars have better parseability than the anti-correlating grammars. This is especially the case for grammars with high $p_\text{branching}$.
